3D Regression Heat Map Analysis of Population Study Data

### 2015-06-23 draft.tex
second round on Bernhards revision; revised paragraphs

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand to assess the model quality.
	%%
  \item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
% %%
% They extend 2D scatter plots to display detailed information about a regression model.
% %%
% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
% %%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow, epidemiological data in general and the SHIP in particular.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many data acquisition modalities available to epidemiologists, which are suitable for different analysis types.
%%
We focus on population studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space.
%%
It also introduces a new information bias, as assumptions are modeled through the discretization.
%%

@@

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
In 1977, Tukey already stated that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
\\\\
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter Space Analysis using Regression Models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} present a conceptual framework of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
The taxonomy allows to sort approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice-views.
%%
We employ the same steps by preprocessing features using correlations with the target and providing a overview over combinations.
%%
M\"{u}hlebacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
\magenta{
%%
Guo et al.~\cite{Guo} presented a multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color coding distances to the line-of-fit.
%%
The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
}
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer} Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

\paragraph{Visual Analysis of Epidemiological Data.}
%%
Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We use their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to depict relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insight into the data.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

Generalized Pairs plots (\texttt{GPLOM'S}) extend the idea of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
\magenta{
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%
}
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical Analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their research agenda acts as a guideline for our design.
%%
Namely, comprising perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al.~\cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
%%
%These subsets are then analyzed in detail, following a drilling-down approach.
%%
They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM'S}) of quality metrics. %describing the features.
%%
%We incorporate the idea of abstracting features for comparison.
%%
Turkay et al.~\cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
The approach was applied to a cognitive aging study, where new hypotheses could be derived.

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
%%
They extract classification rules that serve as basis for our proof-of-concept tests.
%%
Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
%%
\paragraph{Prior Work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%%
%We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%%
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%
We defined an Interactive Visual Analysis workflow for image-centric population study data \cite{Klemm2014VIS}.
%%
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
%%
Information visualizations of population study features were augmented with extracted image data, yielding linked views with both image and non-image information.
%%
The 2D heat map \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
%%
Its great popularity among our domain experts was the inspiration for this work.
%%

@@

Hence, it is associated with step 1 and 2 of the epidemiological workflow presented in Section~\ref{sec:EpidemiologicalWorkflow}.

@@

That is the focus of this work.

@@

Additionally, due to the sensitive nature of population study data, the data has to be handled secure.

@@

Technical measures taken to enable a secure transfer and storage are described later in the Section~\ref{sec:SystemDesign}.

@@

The major advantage of this notation is that it comes natural to anyone familiar with regression analysis, since it is the standard way of expression.

>>>>

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis-generation w.r.t. a target feature using regression models.
	%%
  \item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	%%
  \item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details on demand to assess the model quality.
	%%
  \item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
% %%
% They extend 2D scatter plots to display detailed information about a regression model.
% %%
% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
% %%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow, epidemiological data in general and the SHIP in particular.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
Derived absolute risks indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

We described an Interactive Visual Analysis approach for image-centric cohort study data, which connects to the feature listing step \cite{Klemm2014VIS}.
%%
The methods aim to derive hypotheses through analysis of the data and observations about previously unknown feature correlations.
%%
Employing \emph{hypothesis generation} requires overview visualizations of feature correlations, which are not supported by standard statistical processors.
%%
In this work, we focus on a similar approach to derive insight and even new hypotheses through the data rather than only using it for a confirmatory analysis.
%%
\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
There are many data acquisition modalities available to epidemiologists. %, which are suitable for different analysis types.
%%
We focus on population studies, which impose the largest data sets and yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either regarding specific diseases or with the widest range possible.
%%
The latter allows the data set to be assessed w.r.t. different diseases.
%%
The feature space comprises information about lifestyle, somatometric features, medical parameters, genetic data as well as medical images.
%%
These features are gathered through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow-up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
Medical status features or lifestyle factors are primarily of \emph{dichotomous} (binary) type.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the feature types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space.
%%
It also introduces a new information bias, as assumptions are modeled through the discretization.
%%

@@

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
%It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per participant.
%%
A second cohort, \texttt{SHIP-TREND} was started in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
In 1977, Tukey already stated that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses instead, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} present a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provide a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice-views.
%%
We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlebacher et al. focus on a smaller number of features, which can be assessed in more detail, yielding a plot-matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
\magenta{
%%
Guo et al.~\cite{Guo} presented multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color coding distances to the line-of-fit.
%%
The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
}
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer}, Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

\paragraph{Visual analysis of epidemiological data.}
%%
Zhang et al.~\cite{Zhang2014, Gotz2014} is closest to our work regarding the application of visual analysis on cohort study data.
%%
They present \emph{Cohort Analysis via Visual Analytics} (\emph{CAVA}), a framework that distinguishes three major elements of cohort study data analysis: \emph{cohort data} (and its manipulation using operations), \emph{views} and \emph{analytics}.
%%
They use the system to find longitudinal pathways for diseases on the basis of health records.
%%
Sub-groups are automatically created by incorporating \emph{analytics} and through selection in expert-guided \emph{views}.
%%
%They do not go into the detail w.r.t the analytics used, they seem to not incorporate confounder analysis (other than dividing groups by gender) and the user study lacks concrete goals.
%%
%Their data differs strongly from ours, having multiple moments for each subject, yielding a approach focused on finding pathways of interest.
%%
We use their requirements as guidelines, which incorporate \emph{flexible} and \emph{iterative analysis}. %, meaning that the system has to adapt to different requirements in different analysis steps.
%%
%We implement this by supporting both hypothesis-based and hypothesis-free analysis.
%%
Angelelli et al.~\cite{Angelelli} visualize image-derived an non-image data using cube data structures with focus on comparison and knowledge extraction.
%%
Their proposed engine represents multiple heterogenous cohort study data sets as normalized data cubes.
%%
%Thus, data redundancy is avoided and runtime measure aggregation can be carried out when measures of the different cubes have to be combined.
%%
%This allows for saving memory space.
%%
%Their presented aggregation functions allow for a fast data calculation to map it using different linked visualizations, such as scatter plots, bar charts or histograms.
%%
They use \emph{Pearsson's} $r$ to characterize relationships with target features and employ list views and scatter plots to visualize and rank them.
%%
%We use regression models and associated $R^2$ values to depict relationships between a target feature and multiple independent features instead of one and employ the information into one global visualization.
%%
In contrast, we aim for a fast large scale correlation analysis incorporating many features to derive insight into the data.
%%
%\paragraph{Visual Analysis in Public Health.}
%%--
% Shneiderman et al.~\cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data. %, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
% %%
% %They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
% %%
% %We focus on the latter.
% %%
% %Shneiderman and colleagues also present challenges,
% The presented challenges in this application require the systems to be sufficiently comprehensible and intuitive to fit into the short time span that clinicians have for analyzing data.
% %%
% Furthermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
%We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.
%%

@@

Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
\magenta{
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%
}
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their research agenda acts as a guideline for our design.
%%
Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al.~\cite{Albuquerque} present an interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
%%
%These subsets are then analyzed in detail, following a drilling-down approach.
%%
They analyze subsets using a drilling-down approach by incorporating scatter plot matrices (\texttt{SPLOM'S}) of quality metrics. %describing the features.
%%
%We incorporate the idea of abstracting features for comparison.
%%
Turkay et al.~\cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
%The approach was applied to a cognitive aging study, where new hypotheses could be derived.

%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al.~\cite{Niemann2014} investigate risk factors of hepatic steatosis using decision trees with interactive data mining tools.
%%
They extract classification rules that serve as basis for our proof-of-concept tests.
%%
Niemann et al.~\cite{Niemann2015} improved the classification performance by generating features (called \emph{evolution features}) that describe latent temporal information across the study waves.
%%
We try to reproduce their results and investigate findings presented by Niemann et al.~\cite{Niemann2014} further.
%%
\paragraph{Prior work.}
In our prior work, we analyzed the healthy aging process of the lumbar spine.
%%
%We utilized semi-automatic algorithms do detect the lumbar spine shape \cite{Klemm2013VMV, Rak2013}.
%%
%The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%
We defined an Interactive Visual Analysis workflow for image-centric population study data \cite{Klemm2014VIS}.
%%
We extended the feature selection step (recall Sec. \ref{sec:Background}) with an iterative analysis loop incorporating group selection and visualization using expert input as well as clustering methods.
%%
Information visualizations of population study features were augmented with extracted image data, yielding linked views with both image and non-image information.
%%
The 2D heat map \cite{Klemm2014VIS} highlights the pairwise correlation between features without a target disease by depicting the \emph{Cram\'{e}r's V} contingency values.
%%
Its great popularity among our domain experts was the inspiration for this work.
%%

@@

Hence, it is associated with step 1 and 2 of the epidemiological workflow (recall Section~\ref{sec:EpidemiologicalWorkflow}).

@@

% That is the focus of this work.

@@

Additionally, due to the sensitive nature of population study data, the data has to be handled securely.

@@

Technical measures to enable a secure transfer and storage are described in Section~\ref{sec:SystemDesign}.

@@

This notation comes natural to anyone familiar with regression analysis, since it is the standard way of expression.
